---
title: "Assignment #6"
author: "Qianyi Yu"
date: "11/16/2024"
output: html_document
---

```{r}
library(fpp)
library(fpp2)

#Import the data
mydata <- read.csv("/Users/angela1230/Desktop/Walmart_Sales.csv")

#Convert the date into correct format
mydata$Date <- as.Date(mydata$Date, format="%d-%m-%Y")

#Filter the data for only store 1
store1_sales <- mydata[mydata$Store == 1, ]

#Sort the data in ascending order
store1_sales <- store1_sales[order(store1_sales$Date, decreasing = FALSE), ]

plot(store1_sales$Temperature, store1_sales$Weekly_Sales, pch=19, col="blue", xlab="Temperature", ylab="Weekly Sales")
plot(store1_sales$Fuel_Price, store1_sales$Weekly_Sales, pch=19, col="blue", xlab="Fuel Price", ylab="Weekly Sales")
plot(store1_sales$CPI, store1_sales$Weekly_Sales, pch=19, col="blue", xlab="CPI", ylab="Weekly Sales")
plot(store1_sales$Unemployment, store1_sales$Weekly_Sales, pch=19, col="blue", xlab="Unemployment", ylab="Weekly Sales")

#perform regression model
lm_temp <- lm(store1_sales$Weekly_Sales ~ store1_sales$Temperature)
lm_fuel <- lm(store1_sales$Weekly_Sales ~ store1_sales$Fuel_Price)
lm_cpi <- lm(store1_sales$Weekly_Sales ~ store1_sales$CPI)
lm_unemp <- lm(store1_sales$Weekly_Sales ~ store1_sales$Unemployment)

#plot regression lines
plot(store1_sales$Temperature, store1_sales$Weekly_Sales, pch=19, col="blue", xlab="Temperature", ylab="Weekly Sales")
lines(store1_sales$Temperature, lm_temp$fitted.values, col="red", lwd=3)

plot(store1_sales$Fuel_Price, store1_sales$Weekly_Sales, pch=19, col="blue", xlab="Fuel Price", ylab="Weekly Sales")
lines(store1_sales$Fuel_Price, lm_fuel$fitted.values, col="red", lwd=3)

plot(store1_sales$CPI, store1_sales$Weekly_Sales, pch=19, col="blue", xlab="CPI", ylab="Weekly Sales")
lines(store1_sales$CPI, lm_cpi$fitted.values, col="red", lwd=3)

plot(store1_sales$Unemployment, store1_sales$Weekly_Sales, pch=19, col="blue", xlab="Unemployment", ylab="Weekly Sales")
lines(store1_sales$Unemployment, lm_unemp$fitted.values, col="red", lwd=3)

#Model Acceptance Steps
summary(lm_temp)
summary(lm_fuel)
summary(lm_cpi)
summary(lm_unemp)
```

#### Based on the R-square values and the p-values, CPI appears to be the best predictor for the weekly sales for walmart store 1. CPI has the highest r-square value of 5.08% and a significant p-value of 0.00679. 

```{r}
#Residual Analysis for CPI
par(mfrow=c(2, 2))
plot(lm_cpi)
smoothScatter(store1_sales$CPI, lm_cpi$residuals, xlab="CPI", ylab="Residuals", main="Residuals vs CPI")
abline(h = 0, col = "red", lwd = 2)
```

#### The CPI predicator shows weak explaination power, the r-square value is 5% showing a limited ability to predict weekly sales. The residual analysis shows issues with non-linearity, heteroscedasticity, and deviations from normality, suggesting that the model does not fully capture the underlying relationship. Some points could affecting the model's accuracy. Addressing these issues through data transformation or exploring non-linear modeling could potentially improve the model's performance.
  